---
title: "Machine Learning Workshop in R - Practice time!"
author: "Azka Javaid, Caleb Ki & Muling Si"
date: "3/20/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, echo=FALSE}
#install.packages("caret")
#install.packages("gee")
#install.packages("ROCR")
require(mosaic)
library(mosaic)
library(class)
library(mosaicData)
library(gee)
library(quantreg)
library(nlme)
library(lme4)
library(mdsr)
library(caret) #loads varImp function 
library(rpart) #defining decision trees
library(partykit) #plot decision tree
library(ROCR) #plots ROC curve 
library(randomForest) 
options(digits=3)
trellis.par.set(theme=theme.mosaic())
```

###K-NN Practice

We use the `iris` data set for this practice.

```{r}
data(iris)
names(iris)
```

There are five variables in this dataset. Try fit knn to predict `Species` using the other four variables.


```{r}
X <- iris[, 1:4]
Y <- iris[, 5]
# Speciesknn <- knn()
# confusion <- tally()
# confusion
# sum(diag(confusion))/nrow(testX)
knn_error_rate <- function(x,y, numNeighbors, z=x){
  y_hat <- knn(train =x, test=z, cl=y, k=numNeighbors)
  return(sum(y_hat!=y)/nrow(x))
}
ks <- c(1:15,20,30,40,50)
#train_rate <- sapply(ks, FUN=knn_error_rate, x=, y=)
knn_error_rates <- data.frame(k=ks, train_rates=train_rate)
ggplot(data=knn_error_rates, aes(x=k,y=train_rate))+geom_point()+geom_line()+ylab("misclassification rate")
```

Partition data in training and test sets by specifying the constant beta below. 

```{r}
beta <- ?
n <- nrow(census)
test <- sample.int(n, size = round(beta * n))
train <- census[-test, ]
test <- census[test, ]
tally(~income, data = train, format = "percent")
```

###Decision Tree Practice 

Make a decision tree model to predict species against variables of interest. 

```{r}
form <- as.formula("Species ~ ")
mod_tree <- rpart(form, data = iris)
```

Plot the decision tree and compute variable importance. 

```{r}
#Hint: use as.party to convert tree to a party object and then plot it
```

Compute model accuracy below by filling in the code below.  

```{r}
pred <- predict(?, test, type = "class")
conf <- table(?, pred)
sum(diag(?))/sum(?) #accuracy 
```

### Practice using clustering


```{r}
require(Ecdat)
data(Housing)
colnames(Housing)
```

Cluster houses based on number of bedrooms, number of bathrooms, and the lotsize of the house with 6 different clusters. Write a description of each of the clusters.

```{r}
Housing <- Housing %>% select(?)
clusts <- Housing %>%
  kmeans(centers = ?) %>%
  fitted("classes") %>%
  as.character()

Housing <- Housing %>% mutate(?)
```
